### 2017-11-02 submission/manuscript.Rmd
change discussion grammar and add conclusion

The driver-passenger model of the microbial role in CRC, as summarized by Flynn [@flynn_metabolic_2016], can be supported with this data for carcinoma but not necessarily for adenoma. The drasitically increased relative risk of disease when considering the CRC associated genera is highly supportive of this type of process. In a driver-passenger scenario it is possible that simply having the driver present or only identifying the passenger is a good enough proxy that the event is occuring. This would account for the observation that there is no synergistic incrase in relative risk when accounting for either the total number or increasing abundance of these genera. The initial establishment of the driver within the system is also dependent on the community that is present and this could explain why adding the community context to our models increases the prediction.

Our carcinoma observations fit the driver-passenger model and support this concept within the framework of the transition from adenoma to carcinoma. In contrast, the adenoma observations do not fit well with this model and suggests that the transition from control to adenoma do not fit this framework. The stepwise decrease in diversity suggests that the adenoma community is not normal but this change is subtle. Although there may be localized changes that do depend on the driver-passenger model, our observations show that they do not involve the CRC associated genera. It is possible to hypothesize that at early stages of the diease how the host interacts to these subtle changes could be the catlyst that causes adenoma formation. Subsequent transition to carcinoma could then fit into the proposed driver-passenger model framework.

Although there are still questions that need to be answered for the microbiota and carcinoma a clearer framework is beginning to develop as to how this occurs. The role of the microbiota in adenoma is still not clear and part of the reason may be because many studies are not powered effectely to observe the small changes reported here. It is realistic to suspect that many changes in carcinoma could easily result in effect sizes that are 30% or more between the case and control. Most of the studies have sufficient power to detect these changes. In contrast, our data suggests that the adenoma effect size is relatively small. None of the studies analyzed  were properly powered to detect a 10% or lower change between case and controls and this may well be the range in which differences occur for adenoma. Future studies investigating adenoma and the microbiota need to take these factors into consideration if we are to work out the role of the microbiota in adenoma genesis.

@@

-aggregate studies done with feces and stool to investigate the role of the microbiota in adenoma and carcinoma.

- Significantly adds to the existing research using this approach by expanding both the total individuals included and the scope.

- leads to clear identifications of the challenges that lie ahead with respect to adenoma

>>>>

The driver-passenger model of the microbial role in CRC, as summarized by Flynn [@flynn_metabolic_2016], can be supported with this data for carcinoma but not necessarily for adenoma. The drasitically increased relative risk of disease when considering the CRC associated genera is highly supportive of this type of process. In a driver-passenger scenario it is possible that simply having the driver present or only identifying the passenger is a good enough proxy that the event is occuring. This would account for the observation that there is no synergistic increase in relative risk when accounting for either the total number or increasing abundance of these genera. The initial establishment of the driver within the system is also dependent on the community that is present and this is supported by the observation that when adding the community context to our models along with the CRC associated genera the model AUC increases.

Our carcinoma observations fit the driver-passenger model and support this concept within the framework of the transition from adenoma to carcinoma. In contrast, the adenoma observations do not fit well with this model and suggests that the transition from control to adenoma do not fit this framework. The stepwise decrease in diversity suggests that the adenoma community is not normal but this change is subtle. Although there may be localized changes that do depend on the driver-passenger model, our observations show that they do not involve the CRC associated genera. It is possible to hypothesize that at early stages of the diease, how the host interacts to these subtle changes could be the catlyst that causes adenoma formation. Subsequent transition to carcinoma could then fit into the proposed driver-passenger model framework.

Although there are still questions that need to be answered for the microbiota and carcinoma a clearer framework is beginning to develop as to how this occurs. The role of the microbiota in adenoma is still not clear and part of the reason may be because many studies are not powered effectely to observe the small changes reported here. It is realistic to suspect that many changes in carcinoma could easily result in effect sizes that are 30% or more between the case and control. Most of the studies analyzed have sufficient power to detect these changes. In contrast, our data suggests that the adenoma effect size is relatively small. None of the studies analyzed were properly powered to detect a 10% or lower change between case and controls and this may well be the range in which differences occur in adenoma. Future studies investigating adenoma and the microbiota need to take these factors into consideration if we are to work out the role of the microbiota in adenoma genesis.

@@

By aggregating together a large collection of studies from both feces and tissue we are able to provide information in support of the driver-passenger model in the context of carcinoma. However, within the context of adenoma it is less clear that this relationship exists. These observations highlight the importance of power and sample number considerations when considering investigations into the microbiota and adenoma due to the subtle changes in community. This study helps to identify the problems that have been solved and the challenges that lie ahead in the invesitgation of colorectal cancer and the microbiota.

### 2017-11-02 submission/manuscript.tex
change discussion grammar and add conclusion

The driver-passenger model of the microbial role in CRC, as summarized
by Flynn {[}1{]}, can be supported with this data for carcinoma but not
necessarily for adenoma. The drasitically increased relative risk of
disease when considering the CRC associated genera is highly supportive
of this type of process. In a driver-passenger scenario it is possible
that simply having the driver present or only identifying the passenger
is a good enough proxy that the event is occuring. This would account
for the observation that there is no synergistic incrase in relative
risk when accounting for either the total number or increasing abundance
of these genera. The initial establishment of the driver within the
system is also dependent on the community that is present and this could
explain why adding the community context to our models increases the
prediction.

Our carcinoma observations fit the driver-passenger model and support
this concept within the framework of the transition from adenoma to
carcinoma. In contrast, the adenoma observations do not fit well with
this model and suggests that the transition from control to adenoma do
not fit this framework. The stepwise decrease in diversity suggests that
the adenoma community is not normal but this change is subtle. Although
there may be localized changes that do depend on the driver-passenger
model, our observations show that they do not involve the CRC associated
genera. It is possible to hypothesize that at early stages of the diease
how the host interacts to these subtle changes could be the catlyst that
causes adenoma formation. Subsequent transition to carcinoma could then
fit into the proposed driver-passenger model framework.

Although there are still questions that need to be answered for the
microbiota and carcinoma a clearer framework is beginning to develop as
to how this occurs. The role of the microbiota in adenoma is still not
clear and part of the reason may be because many studies are not powered
effectely to observe the small changes reported here. It is realistic to
suspect that many changes in carcinoma could easily result in effect
sizes that are 30\% or more between the case and control. Most of the
studies have sufficient power to detect these changes. In contrast, our
data suggests that the adenoma effect size is relatively small. None of
the studies analyzed were properly powered to detect a 10\% or lower
change between case and controls and this may well be the range in which
differences occur for adenoma. Future studies investigating adenoma and
the microbiota need to take these factors into consideration if we are
to work out the role of the microbiota in adenoma genesis.

@@

-aggregate studies done with feces and stool to investigate the role of
the microbiota in adenoma and carcinoma.

\begin{itemize}
\item
  Significantly adds to the existing research using this approach by
  expanding both the total individuals included and the scope.
\item
  leads to clear identifications of the challenges that lie ahead with
  respect to adenoma
\end{itemize}

>>>>

The driver-passenger model of the microbial role in CRC, as summarized
by Flynn {[}1{]}, can be supported with this data for carcinoma but not
necessarily for adenoma. The drasitically increased relative risk of
disease when considering the CRC associated genera is highly supportive
of this type of process. In a driver-passenger scenario it is possible
that simply having the driver present or only identifying the passenger
is a good enough proxy that the event is occuring. This would account
for the observation that there is no synergistic increase in relative
risk when accounting for either the total number or increasing abundance
of these genera. The initial establishment of the driver within the
system is also dependent on the community that is present and this is
supported by the observation that when adding the community context to
our models along with the CRC associated genera the model AUC increases.

Our carcinoma observations fit the driver-passenger model and support
this concept within the framework of the transition from adenoma to
carcinoma. In contrast, the adenoma observations do not fit well with
this model and suggests that the transition from control to adenoma do
not fit this framework. The stepwise decrease in diversity suggests that
the adenoma community is not normal but this change is subtle. Although
there may be localized changes that do depend on the driver-passenger
model, our observations show that they do not involve the CRC associated
genera. It is possible to hypothesize that at early stages of the
diease, how the host interacts to these subtle changes could be the
catlyst that causes adenoma formation. Subsequent transition to
carcinoma could then fit into the proposed driver-passenger model
framework.

Although there are still questions that need to be answered for the
microbiota and carcinoma a clearer framework is beginning to develop as
to how this occurs. The role of the microbiota in adenoma is still not
clear and part of the reason may be because many studies are not powered
effectely to observe the small changes reported here. It is realistic to
suspect that many changes in carcinoma could easily result in effect
sizes that are 30\% or more between the case and control. Most of the
studies analyzed have sufficient power to detect these changes. In
contrast, our data suggests that the adenoma effect size is relatively
small. None of the studies analyzed were properly powered to detect a
10\% or lower change between case and controls and this may well be the
range in which differences occur in adenoma. Future studies
investigating adenoma and the microbiota need to take these factors into
consideration if we are to work out the role of the microbiota in
adenoma genesis.

@@

By aggregating together a large collection of studies from both feces
and tissue we are able to provide information in support of the
driver-passenger model in the context of carcinoma. However, within the
context of adenoma it is less clear that this relationship exists. These
observations highlight the importance of power and sample number
considerations when considering investigations into the microbiota and
adenoma due to the subtle changes in community. This study helps to
identify the problems that have been solved and the challenges that lie
ahead in the invesitgation of colorectal cancer and the microbiota.
